Reducing Linearizability of Priority Queues and More Data Structures to State Reachability

### 2017-04-18 appendix-EPQ.tex
change appendix

Let us model extended priority queue as an LTS $\textit{LTS}_e = (Q,\Sigma,\rightarrow,q_0)$ as follows:

@@

    \item[-] $q_1 \xrightarrow{\textit{rm}(a)} q_2$, if $q_1$ maps $p$ into $l \cdot a$ for some finite sequence $l$, and $q_2$ is the same as $q_1$, except for $p$, where it maps $p$ into $l$. We also require that for each priority $p'$ such that $p' <_{\mathbb{P}} p$, $q_1$ and $q_2$ map $p'$ into $\epsilon$.

@@

A path of an LTS is a finite transition sequence $q_0\xrightarrow{\beta_1}q_1\overset{\beta_2}{\longrightarrow}\ldots\overset{\beta_k}{\longrightarrow}q_k$ for $k\geq 0$, where $q_0$ is the initial state of the LTS. A trace of an LTS is a finite sequence $\beta_1 \cdot \beta_2 \cdot \ldots \cdot \beta_k$, where $k \geq 0$ if there exists a path $q_0\overset{\beta_1}{\longrightarrow}q_1\overset{\beta_2}{\longrightarrow}\ldots\overset{\beta_k}{\longrightarrow}q_k$ of the LTS. Let $\textit{EPQ}_s$ be the set of traces of $\textit{LTS}_e$. The following lemma states that the language generated by our rules equals the set of traces of $\textit{EPQ}_s$.

@@

To prove that $\textit{EPQ} \subseteq \textit{EPQ}_s$, we prove that each sequence in $\textit{EPQ}$ is also in $\textit{EPQ}_s$ by induction:

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] It is obvious that $\epsilon \in \textit{EPQ}_s$.

\item[-] If $l_1 \in \textit{EPQ}_s$ and $l_1 \xrightarrow{\textit{EPQ}_1} l_2$. Then we need to prove that $l_2 \in \textit{EPQ}_s$. Let $l_1 = u \cdot v \cdot w$, such that $\textit{Guard}(u,v,w,\textit{item},\textit{pri})$ of $\textit{EPQ}_1$ holds, and $l_2 = u \cdot \textit{put}(\textit{itm},\textit{pri}) \cdot v \cdot \textit{rm}(\textit{itm})$.

    Assume that $u = \alpha_1 \cdot \ldots \cdot \alpha_i$, $v = \alpha_{\textit{i+1}} \cdot \ldots \cdot \alpha_j$ and $w = \alpha_{\textit{j+1}} \cdot \ldots \cdot \alpha_m$. Assume that $q_0 \xrightarrow{\alpha_1} q_1 \ldots \xrightarrow{\alpha_i} q_i \xrightarrow{\alpha_{\textit{i+1}}} q_{\textit{i+1}} \ldots  \xrightarrow{\alpha_j} q_j \xrightarrow{\alpha_{\textit{j+1}}} q_{\textit{j+1}} \ldots \xrightarrow{\alpha_m} q_m$ is the path of $l_1$ on $\textit{EPQ}_s$. For each $i \leq k \leq j$, let $q'_k$ be the same as $q_k$, except that $q'_k$ maps $\textit{pri}$ into $\textit{item} \cdot l_k$ and $q_k$ maps $\textit{pri}$ into $l_k$ for some finite sequence $l_k$.

    We already know that $q_0 \xrightarrow{\alpha_1} q_1 \ldots \xrightarrow{\alpha_i} q_i$, and it is obvious that $q_i \xrightarrow{\textit{put}(\textit{itm},\textit{pri})} q'_i$. Since (1) all $\textit{put}$ with priority $\textit{pri}$ is in $u$, and (2) in $u \cdot v$, only items with priority either incomparable, or less, or equal than $\textit{pri}$ is removed, we can see that it is safe to add to each $q_k$ ($1 \leq k \leq j$) with a newest $\textit{itm}$ with priority $\textit{pri}$. Or we can say, $q'_i \xrightarrow{\alpha_{\textit{i+1}}} q'_{\textit{i+1}} \ldots \xrightarrow{\alpha_j} q'_j$ are transitions of $\textit{LTS}_e$. Since $\textit{matched-C}(u \cdot v)$ holds, we can see that $q_j$ maps each priority that is smaller than $\textit{pri}$ into $\epsilon$ and maps $\textit{pri}$ into $\epsilon$, and $q'_j$ maps each priority that is smaller than $\textit{pri}$ into $\epsilon$ and maps $\textit{pri}$ into $\textit{itm}$. Then, we can see that $q'_j \xrightarrow{\textit{rm}(\textit{itm})} q_j$. We already know that that $q_j \xrightarrow{\alpha_{\textit{j+1}}} q_{\textit{j+1}} \ldots \xrightarrow{\alpha_m} q_m$. Therefore, we can see that $l_2 = u \cdot \textit{put}(\textit{itm},\textit{pri}) \cdot v \cdot \textit{rm}(\textit{itm}) \in \textit{EPQ}_s$.

\item[-] If $l_1 \in \textit{EPQ}_s$ and $l_1 \xrightarrow{\textit{EPQ}_2} l_2$. Then we need to prove that $l_2 \in \textit{EPQ}_s$. Let $l_1 = u \cdot v$, such that $\textit{Guard}(u,v,\textit{item},\textit{pri})$ of $\textit{EPQ}_2$ holds, and $l_2 = u \cdot \textit{put}(\textit{itm},\textit{pri}) \cdot v$.

    Assume that $u = \alpha_1 \cdot \ldots \cdot \alpha_i$ and $v = \alpha_{\textit{i+1}} \cdot \ldots \cdot \alpha_m$. Assume that $q_0 \xrightarrow{\alpha_1} q_1 \ldots \xrightarrow{\alpha_i} q_i \xrightarrow{\alpha_{\textit{i+1}}} q_{\textit{i+1}} \ldots  \xrightarrow{\alpha_m} q_m$ is the path of $l_1$ on $\textit{EPQ}_s$. For each $i \leq k \leq m$, let $q'_k$ be the same as $q_k$, except that $q'_k$ maps $\textit{pri}$ into $\textit{item} \cdot l_k$ and $q_k$ maps $\textit{pri}$ into $l_k$ for some finite sequence $l_k$.

    We already know that $q_0 \xrightarrow{\alpha_1} q_1 \ldots \xrightarrow{\alpha_i} q_i$, and it is obvious that$q_i \xrightarrow{\textit{put}(\textit{itm},\textit{pri})} q'_i$. Since (1) all $\textit{put}$ with priority $\textit{pri}$ is in $u$, (2) in $u \cdot v$, only items with priority either incomparable, or less, or equal than $\textit{pri}$ is removed, we can see that it is safe to add to each $q_k$ ($1 \leq k \leq m$) with a newest $\textit{itm}$ with priority $\textit{pri}$. Or we can say, $q'_i \xrightarrow{\alpha_{\textit{i+1}}} q'_{\textit{i+1}} \ldots \xrightarrow{\alpha_m} q'_m$ are transitions of $\textit{LTS}_e$. Therefore, we can see that $l_2 = u \cdot \textit{put}(\textit{itm},\textit{pri}) \cdot v \in \textit{EPQ}_s$.

\item[-] If $l_1 \in \textit{EPQ}_s$ and $l_1 \xrightarrow{\textit{EPQ}_3} l_2$. Then we need to prove that $l_2 \in \textit{EPQ}_s$. Let $l_1 = u \cdot v$, such that $\textit{Guard}(u,v)$ of $\textit{EPQ}_3$ holds, and $l_2 = u \cdot \textit{rm}(\textit{empty}) \cdot v$.

    Assume that $u = \alpha_1 \cdot \ldots \cdot \alpha_i$ and $v = \alpha_{\textit{i+1}} \cdot \ldots \cdot \alpha_m$. Assume that $q_0 \xrightarrow{\alpha_1} q_1 \ldots \xrightarrow{\alpha_i} q_i \xrightarrow{\alpha_{\textit{i+1}}} q_{\textit{i+1}} \ldots  \xrightarrow{\alpha_m} q_m$ is the path of $l_1$ on $\textit{EPQ}_s$.

    We already know that $q_0 \xrightarrow{\alpha_1} q_1 \ldots \xrightarrow{\alpha_i} q_i$. Since $\textit{matched-All}(u)$ holds, we can see that $q_i$ maps each element in $\mathbb{P}$ into $\epsilon$, and then $q_i \xrightarrow{\textit{rm}(\textit{empty})} q_i$. We already know that $q_i \xrightarrow{\alpha_{\textit{i+1}}} q_{\textit{i+1}} \ldots \xrightarrow{\alpha_m} q_m$. Therefore, we can see that $l_2 = u \cdot \textit{rm}(\textit{empty}) \cdot v \in \textit{EPQ}_s$.
\end{itemize}

To prove that $\textit{EPQ}_s \subseteq \textit{EPQ}$, we show that given $l_2 \in \textit{EPQ}_s$ and $l_2 \neq \epsilon$, how to construct a sequence $l_1$, such that $l_1 \xrightarrow{R} l_2$ for some rule $R$, and $l_1 \in \textit{EPQ}_s$. Based on this, we can decompose a sequence of $\textit{EPQ}_s$ into $\epsilon$, while the reverse process is the reason of why this sequence is in $\textit{EPQ}$. Note that from a $l_2$ we may construct more than one $l_1$, and this does not influence the correctness of our proof.

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] If $l_2$ contains $\textit{rm}(\textit{empty})$: Assume that $l_2 = u \cdot \textit{rm}(\textit{empty}) \cdot v$. It is easy to see that $\textit{matched-All}(u)$ holds. Let $l_1 = u \cdot v$. It is easy to see that $l_1 \xrightarrow{\textit{EPQ}_3} l_2$, and $l_1$ satisfy the guard of $\textit{EPQ}_3$.

    Assume that $u = \alpha_1 \cdot \ldots \cdot \alpha_i$ and $v = \alpha_{\textit{i+1}} \cdot \ldots \cdot \alpha_m$. Since We already know that $q_0 \xrightarrow{\alpha_1} q_1 \ldots \xrightarrow{\alpha_i} q_i \xrightarrow{\textit{rm}(\textit{empty})} q'_i \xrightarrow{\alpha_{\textit{i+1}}} q_{\textit{i+1}} \ldots  \xrightarrow{\alpha_m} q_m$ is transitions of $\textit{LTS}_e$. Since $\textit{matched-All}(u)$ holds, we can see that $q_i = q'_i$, and they map each element in $\mathbb{P}$ into $\epsilon$. Then we can see that $q_0 \xrightarrow{\alpha_1} q_1 \ldots \xrightarrow{\alpha_i} q_i \xrightarrow{\alpha_{\textit{i+1}}} q_{\textit{i+1}} \ldots  \xrightarrow{\alpha_m} q_m$ is transitions of $\textit{LTS}_e$, and $l_1 \in \textit{EPQ}_s$.

\item[-] If $l_2$ does not contain $\textit{rm}(\textit{empty})$, $\textit{pri}$ is one of maximal priority of $l_2$, and items in $l_2$ with priority $\textit{pri}$ are unmatched $\textit{put}$ and (possibly) matched $\textit{put}$:

    Assume that $l_2 = u \cdot \textit{put}(\textit{itm},\textit{pri}) \cdot v$, such that all $\textit{put}$ with priority $\textit{pri}$ of $u \cdot v$ is in $u$. Let $l_1 = u \cdot v$. According to construction of $\textit{LTS}_e$, we can see that $l_1$ satisfies the guard of $\textit{EPQ}_2$, and $l_1 \xrightarrow{\textit{EPQ}_2} l_2$.

    Assume that $u = \alpha_1 \cdot \ldots \cdot \alpha_i$ and $v = \alpha_{\textit{i+1}} \cdot \ldots \cdot \alpha_m$. We already know that $\textit{pa} = q_0 \xrightarrow{\alpha_1} q_1 \ldots \xrightarrow{\alpha_i} q_i \xrightarrow{\textit{put}(\textit{itm},\textit{pri})} q_{\textit{i+}} \xrightarrow{\alpha_{\textit{i+1}}} q_{\textit{i+1}} \ldots \xrightarrow{\alpha_m} q_m$ are transitions of $\textit{LTS}_e$. For each $\textit{i+1} \leq k \leq m$, let $q'_k$ be the same as $q_k$, except that $q_k$ maps $\textit{pri}$ into some $\textit{itm} \cdot l_k$ for some finite sequence $l_k$, and $q'_k$ maps $\textit{pri}$ into $l_k$. Since (1) all $\textit{put}$ with priority $\textit{pri}$ of $u \cdot v$ is in $u$ and (2) $\textit{pri}$ is one of maximal priority of $l_2$, it is safe to remove $\textit{itm}$ without influence other transitions of $\textit{pa}$. Or we can say, $q_0 \xrightarrow{\alpha_1} q_1 \ldots \xrightarrow{\alpha_i} q_i \xrightarrow{\alpha_{\textit{i+1}}} q'_{\textit{i+1}} \ldots \xrightarrow{\alpha_m} q'_m$ are transitions of $\textit{LTS}_e$. Therefore, $l_1 \in \textit{EPQ}_s$.

\item[-] If $l_2$ does not contain $\textit{rm}(\textit{empty})$, $\textit{pri}$ is one of maximal priority of $l_2$, and items in $l_2$ with priority $\textit{pri}$ are matched $\textit{put}$:

    Assume that $l_2 = u \cdot \textit{put}(\textit{itm},\textit{pri}) \cdot v \cdot \textit{rm}(\textit{itm}) \cdot w$, such that all $\textit{put}$ with priority $\textit{pri}$ of $u \cdot v \cdot w$ is in $u$. Let $l_1 = u \cdot v \cdot w$. According to construction of $\textit{LTS}_e$, we can see that $l_1$ satisfies the guard of $\textit{EPQ}_1$. We can also see that $l_1 \xrightarrow{\textit{EPQ}_1} l_2$.

    Assume that $u = \alpha_1 \cdot \ldots \cdot \alpha_i$, $v = \alpha_{\textit{i+1}} \cdot \ldots \cdot \alpha_j$ and $w = \alpha_{\textit{j+1}} \cdot \ldots \cdot \alpha_m$. We already know that $q_0 \xrightarrow{\alpha_1} q_1 \ldots \xrightarrow{\alpha_i} q_i \xrightarrow{\textit{put}(\textit{itm},\textit{pri})} q_{\textit{i+}} \xrightarrow{\alpha_{\textit{i+1}}} q_{\textit{i+1}} \ldots \xrightarrow{\alpha_j} q_j \xrightarrow{\textit{rm}(\textit{itm})} q_{\textit{j+}} \xrightarrow{\alpha_{\textit{j+1}}} q_{\textit{j+1}} \ldots \xrightarrow{\alpha_m} q_m$. For each $\textit{i+1} \leq k \leq j$, let $q'_k$ be the same as $q_k$, except that $q_k$ maps $\textit{pri}$ into $\textit{itm} \cdot l_k$ for some finite sequence $l_k$, and $q'_k$ maps $\textit{pri}$ into $l_k$. Since (1) $\textit{pri}$ is one of maximal priority in $l_2$, (2) $\textit{itm}$ is the newest item with priority $\textit{pri}$ in $l_2$, and (3) $\textit{itm}$ is not removed until $\textit{rm}(\textit{itm})$, we know that whether we keep $\textit{itm}$ or remove it will not influence transitions from $q_{\textit{i+1}}$ to $q_j$. Then we can see that $q_0 \xrightarrow{\alpha_1} q_1 \ldots \xrightarrow{\alpha_i} q_i \xrightarrow{\alpha_{\textit{i+1}}} q'_{\textit{i+1}} \ldots \xrightarrow{\alpha_j} q'_j \xrightarrow{\alpha_{\textit{j+1}}} q_{\textit{j+1}} \ldots \xrightarrow{\alpha_m} q_m$ are transitions of $\textit{LTS}_e$. Therefore, $l_1 \in \textit{EPQ}_s$.
\end{itemize}

>>>>

Let us model priority queue as an LTS $\textit{LTS}_e = (Q,\Sigma,\rightarrow,q_0)$ as follows:

@@

    \item[-] $q_1 \xrightarrow{\textit{rm}(a)} q_2$, if $q_1$ maps $p$ into $l \cdot a$ for some finite sequence $l$, and $q_2$ is the same as $q_1$, except for $p$, where it maps $p$ into $l$. We also require that for each priority $p'$ such that $p' \prec p$, $q_1$ and $q_2$ map $p'$ into $\epsilon$.

@@

A path of an LTS is a finite transition sequence $q_0\xrightarrow{\beta_1}q_1\overset{\beta_2}{\longrightarrow}\ldots\overset{\beta_k}{\longrightarrow}q_k$ for $k\geq 0$, where $q_0$ is the initial state of the LTS. A trace of an LTS is a finite sequence $\beta_1 \cdot \beta_2 \cdot \ldots \cdot \beta_k$, where $k \geq 0$ if there exists a path $q_0\overset{\beta_1}{\longrightarrow}q_1\overset{\beta_2}{\longrightarrow}\ldots\overset{\beta_k}{\longrightarrow}q_k$ of the LTS. Let $\textit{PQ}_s$ be the set of traces of $\textit{LTS}_e$. The following lemma states that $\textit{PQ} = \textit{PQ}_s$.

@@

To prove that $\textit{PQ} \subseteq \textit{PQ}_s$, we first prove that $\textit{PQ}_{\neq} \subseteq \textit{PQ}_{s\neq}$ by induction.

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] It is obvious that $\epsilon \in \textit{PQ}_s$.

\item[-] If $l_1 \in \textit{PQ}_{s\neq}$ and $l_2 \xrightarrow{\textit{PQ}_1} l_1$. Then we need to prove that $l_2 \in \textit{PQ}_s$. We know that $l_1 = u \cdot v \cdot w$, such that $\textit{PQ}_1(u,v,w,p)$ holds, and $l_2 = u \cdot \textit{put}(\textit{itm},p) \cdot v \cdot \textit{rm}(\textit{itm})$.

    Assume that $u = \alpha_1 \cdot \ldots \cdot \alpha_i$, $v = \alpha_{\textit{i+1}} \cdot \ldots \cdot \alpha_j$ and $w = \alpha_{\textit{j+1}} \cdot \ldots \cdot \alpha_m$. Assume that $q_0 \xrightarrow{\alpha_1} q_1 \ldots \xrightarrow{\alpha_i} q_i \xrightarrow{\alpha_{\textit{i+1}}} q_{\textit{i+1}} \ldots  \xrightarrow{\alpha_j} q_j \xrightarrow{\alpha_{\textit{j+1}}} q_{\textit{j+1}} \ldots \xrightarrow{\alpha_m} q_m$ is the path of $l_1$ on $\textit{PQ}_s$. For each $i \leq k \leq j$, let $q'_k$ be the same as $q_k$, except that $q'_k$ maps $p$ into $\textit{item} \cdot l_k$ and $q_k$ maps $\textit{pri}$ into $l_k$ for some finite sequence $l_k$.

    We already know that $q_0 \xrightarrow{\alpha_1} q_1 \ldots \xrightarrow{\alpha_i} q_i$, and it is obvious that $q_i \xrightarrow{\textit{put}(\textit{itm},p)} q'_i$. Since (1) all $\textit{put}$ with priority $p$ is in $u$, and (2) in $u \cdot v$, only items with priority either incomparable, or less, or equal than $p$ is removed, we can see that it is safe to add to each $q_k$ ($1 \leq k \leq j$) with a newest $\textit{itm}$ with priority $p$. Or we can say, $q'_i \xrightarrow{\alpha_{\textit{i+1}}} q'_{\textit{i+1}} \ldots \xrightarrow{\alpha_j} q'_j$ are transitions of $\textit{LTS}_e$. Since $\textit{matched}_{\prec}(u \cdot v,p)$ holds, we can see that $q_j$ maps each priority that is smaller than $p$ into $\epsilon$ and maps $p$ into $\epsilon$, and $q'_j$ maps each priority that is smaller than $p$ into $\epsilon$ and maps $p$ into $\textit{itm}$. Then, we can see that $q'_j \xrightarrow{\textit{rm}(\textit{itm})} q_j$. We already know that that $q_j \xrightarrow{\alpha_{\textit{j+1}}} q_{\textit{j+1}} \ldots \xrightarrow{\alpha_m} q_m$. Therefore, we can see that $l_2 = u \cdot \textit{put}(\textit{itm},p) \cdot v \cdot \textit{rm}(\textit{itm}) \in \textit{PQ}_s$.

\item[-] If $l_1 \in \textit{PQ}_{s\neq}$ and $l_2 \xrightarrow{\textit{PQ}_2} l_1$. Then we need to prove that $l_2 \in \textit{PQ}_s$. We know that $l_1 = u \cdot v$, such that $\textit{PQ}_2(u,v,p)$ holds, and $l_2 = u \cdot \textit{put}(\textit{itm},p) \cdot v$.

    Assume that $u = \alpha_1 \cdot \ldots \cdot \alpha_i$ and $v = \alpha_{\textit{i+1}} \cdot \ldots \cdot \alpha_m$. Assume that $q_0 \xrightarrow{\alpha_1} q_1 \ldots \xrightarrow{\alpha_i} q_i \xrightarrow{\alpha_{\textit{i+1}}} q_{\textit{i+1}} \ldots  \xrightarrow{\alpha_m} q_m$ is the path of $l_1$ on $\textit{PQ}_s$. For each $i \leq k \leq m$, let $q'_k$ be the same as $q_k$, except that $q'_k$ maps $p$ into $\textit{item} \cdot l_k$ and $q_k$ maps $p$ into $l_k$ for some finite sequence $l_k$.

    We already know that $q_0 \xrightarrow{\alpha_1} q_1 \ldots \xrightarrow{\alpha_i} q_i$, and it is obvious that $q_i \xrightarrow{\textit{put}(\textit{itm},p)} q'_i$. Since (1) all $\textit{put}$ with priority $p$ is in $u$, (2) in $u \cdot v$, only items with priority either incomparable, or less, or equal than $p$ is removed, we can see that it is safe to add to each $q_k$ ($1 \leq k \leq m$) with a newest $\textit{itm}$ with priority $p$. Or we can say, $q'_i \xrightarrow{\alpha_{\textit{i+1}}} q'_{\textit{i+1}} \ldots \xrightarrow{\alpha_m} q'_m$ are transitions of $\textit{LTS}_e$. Therefore, we can see that $l_2 = u \cdot \textit{put}(\textit{itm},p) \cdot v \in \textit{PQ}_s$.

\item[-] If $l_1 \in \textit{PQ}_{s\neq}$ and $l_2 \xrightarrow{\textit{PQ}_3} l_1$. Then we need to prove that $l_2 \in \textit{PQ}_s$. We know that $l_1 = u \cdot v$, such that $\textit{PQ}_3(u,v)$ holds, and $l_2 = u \cdot \textit{rm}(\textit{empty}) \cdot v$.

    Assume that $u = \alpha_1 \cdot \ldots \cdot \alpha_i$ and $v = \alpha_{\textit{i+1}} \cdot \ldots \cdot \alpha_m$. Assume that $q_0 \xrightarrow{\alpha_1} q_1 \ldots \xrightarrow{\alpha_i} q_i \xrightarrow{\alpha_{\textit{i+1}}} q_{\textit{i+1}} \ldots  \xrightarrow{\alpha_m} q_m$ is the path of $l_1$ on $\textit{PQ}_s$.

    We already know that $q_0 \xrightarrow{\alpha_1} q_1 \ldots \xrightarrow{\alpha_i} q_i$. Since $\textit{matched}(u)$ holds, we can see that $q_i$ maps each element in $\mathbb{P}$ into $\epsilon$, and then $q_i \xrightarrow{\textit{rm}(\textit{empty})} q_i$. We already know that $q_i \xrightarrow{\alpha_{\textit{i+1}}} q_{\textit{i+1}} \ldots \xrightarrow{\alpha_m} q_m$. Therefore, we can see that $l_2 = u \cdot \textit{rm}(\textit{empty}) \cdot v \in \textit{PQ}_s$.
\end{itemize}

To prove that $\textit{PQ} \subseteq \textit{PQ}_s$. It is easy to see that $\textit{PQ}$ and $\textit{PQ}_s$ are both data-independent. Given $e_1 \in \textit{PQ}$, by data-independence we know that there is a data-differentiated sequential execution $e' \in \textit{PQ}$ and renaming function $f$, such that $e=f(e')$. By $\textit{PQ}_{\neq} \subseteq \textit{PQ}_{s\neq}$, we know that $e' \in \textit{PQ}_{s\neq}$. By data-independence, we know that $e \in \textit{PQ}_s$.

To prove that $\textit{PQ}_s \subseteq \textit{PQ}$, let us first prove $\textit{PQ}_{s\neq} \subseteq \textit{PQ}_{\neq}$. We show that given $l_2 \in \textit{PQ}_{s\neq}$, how to construct a sequence $l_1$, such that $l_2 \xrightarrow{R} l_1$ for some $R$, and $l_1 \in \textit{PQ}_s$. Based on this, we can decompose a sequence of $\textit{PQ}_s$ into $\epsilon$, and this process ensures that this sequence is in $\textit{PQ}$. Note that from a $l_2$ we may construct more than one $l_1$, and this does not influence the correctness of our proof.

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] If $l_2$ contains $\textit{rm}(\textit{empty})$: Assume that $l_2 = u \cdot \textit{rm}(\textit{empty}) \cdot v$. It is easy to see that $\textit{matched}(u)$ holds. Let $l_1 = u \cdot v$. It is easy to see that $l_2 \xrightarrow{\textit{PQ}_3} l_1$, and $\textit{PQ}_3(u,v)$ holds.

    Assume that $u = \alpha_1 \cdot \ldots \cdot \alpha_i$ and $v = \alpha_{\textit{i+1}} \cdot \ldots \cdot \alpha_m$. Since We already know that $q_0 \xrightarrow{\alpha_1} q_1 \ldots \xrightarrow{\alpha_i} q_i \xrightarrow{\textit{rm}(\textit{empty})} q'_i \xrightarrow{\alpha_{\textit{i+1}}} q_{\textit{i+1}} \ldots  \xrightarrow{\alpha_m} q_m$ is transitions of $\textit{LTS}_e$. It is easy to see that $q_i = q'_i$, and they map each element in $\mathbb{P}$ into $\epsilon$. Then we can see that $q_0 \xrightarrow{\alpha_1} q_1 \ldots \xrightarrow{\alpha_i} q_i \xrightarrow{\alpha_{\textit{i+1}}} q_{\textit{i+1}} \ldots  \xrightarrow{\alpha_m} q_m$ is transitions of $\textit{LTS}_e$, and $l_1 \in \textit{PQ}_s$.

\item[-] If $l_2$ does not contain $\textit{rm}(\textit{empty})$, $p$ is one of maximal priority of $l_2$, and items in $l_2$ with priority $p$ are unmatched $\textit{put}$ and (possibly) matched $\textit{put}$:

    Assume that $l_2 = u \cdot \textit{put}(\textit{itm},p) \cdot v$, such that all $\textit{put}$ with priority $p$ of $u \cdot v$ is in $u$. Let $l_1 = u \cdot v$. According to construction of $\textit{LTS}_e$, we can see that $\textit{PQ}_2(u,v,p)$ holds, and $l_2 \xrightarrow{\textit{PQ}_2} l_1$.

    Assume that $u = \alpha_1 \cdot \ldots \cdot \alpha_i$ and $v = \alpha_{\textit{i+1}} \cdot \ldots \cdot \alpha_m$. We already know that $\textit{pa} = q_0 \xrightarrow{\alpha_1} q_1 \ldots \xrightarrow{\alpha_i} q_i \xrightarrow{\textit{put}(\textit{itm},p)} q_{\textit{i+}} \xrightarrow{\alpha_{\textit{i+1}}} q_{\textit{i+1}} \ldots \xrightarrow{\alpha_m} q_m$ are transitions of $\textit{LTS}_e$. For each $\textit{i+1} \leq k \leq m$, let $q'_k$ be the same as $q_k$, except that $q_k$ maps $p$ into some $\textit{itm} \cdot l_k$ for some finite sequence $l_k$, and $q'_k$ maps $p$ into $l_k$. Since (1) all $\textit{put}$ with priority $p$ of $u \cdot v$ is in $u$ and (2) $p$ is one of maximal priority of $l_2$, it is safe to remove $\textit{itm}$ without influence other transitions of $\textit{pa}$. Or we can say, $q_0 \xrightarrow{\alpha_1} q_1 \ldots \xrightarrow{\alpha_i} q_i \xrightarrow{\alpha_{\textit{i+1}}} q'_{\textit{i+1}} \ldots \xrightarrow{\alpha_m} q'_m$ are transitions of $\textit{LTS}_e$. Therefore, $l_1 \in \textit{PQ}_s$.

\item[-] If $l_2$ does not contain $\textit{rm}(\textit{empty})$, $p$ is one of maximal priority of $l_2$, and items in $l_2$ with priority $p$ are matched $\textit{put}$:

    Assume that $l_2 = u \cdot \textit{put}(\textit{itm},p) \cdot v \cdot \textit{rm}(\textit{itm}) \cdot w$, such that all $\textit{put}$ with priority $p$ of $u \cdot v \cdot w$ is in $u$. Let $l_1 = u \cdot v \cdot w$. According to construction of $\textit{LTS}_e$, we can see that $\textit{PQ}_1(u,v,w,p)$ holds, and $l_2 \xrightarrow{\textit{PQ}_1} l_1$.

    Assume that $u = \alpha_1 \cdot \ldots \cdot \alpha_i$, $v = \alpha_{\textit{i+1}} \cdot \ldots \cdot \alpha_j$ and $w = \alpha_{\textit{j+1}} \cdot \ldots \cdot \alpha_m$. We already know that $q_0 \xrightarrow{\alpha_1} q_1 \ldots \xrightarrow{\alpha_i} q_i \xrightarrow{\textit{put}(\textit{itm},\textit{pri})} q_{\textit{i+}} \xrightarrow{\alpha_{\textit{i+1}}} q_{\textit{i+1}} \ldots \xrightarrow{\alpha_j} q_j \xrightarrow{\textit{rm}(\textit{itm})} q_{\textit{j+}} \xrightarrow{\alpha_{\textit{j+1}}} q_{\textit{j+1}} \ldots \xrightarrow{\alpha_m} q_m$. For each $\textit{i+1} \leq k \leq j$, let $q'_k$ be the same as $q_k$, except that $q_k$ maps $p$ into $\textit{itm} \cdot l_k$ for some finite sequence $l_k$, and $q'_k$ maps $p$ into $l_k$. Since (1) $p$ is one of maximal priority in $l_2$, (2) $\textit{itm}$ is the newest item with priority $p$ in $l_2$, and (3) $\textit{itm}$ is not removed until $\textit{rm}(\textit{itm})$, we know that whether we keep $\textit{itm}$ or remove it will not influence transitions from $q_{\textit{i+1}}$ to $q_j$. Then we can see that $q_0 \xrightarrow{\alpha_1} q_1 \ldots \xrightarrow{\alpha_i} q_i \xrightarrow{\alpha_{\textit{i+1}}} q'_{\textit{i+1}} \ldots \xrightarrow{\alpha_j} q'_j \xrightarrow{\alpha_{\textit{j+1}}} q_{\textit{j+1}} \ldots \xrightarrow{\alpha_m} q_m$ are transitions of $\textit{LTS}_e$. Therefore, $l_1 \in \textit{PQ}_s$.
\end{itemize}

To prove that $\textit{PQ}_s \subseteq \textit{PQ}$. Given $e_1 \in \textit{PQ}_s$, by data-independence we know that there is a data-differentiated sequential execution $e' \in \textit{PQ}_s$ and renaming function $f$, such that $e=f(e')$. By $\textit{PQ}_{s\neq} \subseteq \textit{PQ}_{\neq}$, we know that $e' \in \textit{PQ}_{\neq}$. By data-independence, we know that $e \in \textit{PQ}$. \qed
\end {proof}
